Naive Bayes

### 2018-12-18 Naive_bayes.Rmd
update

#Demonstration of Applying Naive Bayes Method

The objective of this project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

@@

P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)

@@

##Loading Data Into R.

```{r}
sms_raw <- read.table("~/Dropbox/Coding_Tools/R_Environment/R_Projects/ta_comments_FLC/data/Bayes/SMSSpamCollection.txt", header = FALSE, sep = "\t", quote = "", col.names = c("type","text"), stringsAsFactors = FALSE)
```

The data frame has 5574 observations with 2 observations of either spam or ham.

@@

Now the "type" variable is a factor with 2 levels. Of the 5574 messages, 747 are spam, which is about 13.4%.

##Text Mining

For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.

```{r}
#install.packages("tm")
library(tm)
```

@@

```{r}
as.character(sms_corpus[[3]])

@@

```{r}
#install.packages("SnowballC")
library(SnowballC)

sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
```

@@

sms_test_labels
      ham      spam 
0.8694706 0.1305294

@@

```{r}
library(wordcloud)
wordcloud(sms_corpus_clean, max.words = 50, random.order = FALSE)
```

Compare wordclouds between spam and ham.
```{r}
spam <- subset(sms_raw, type == "spam")
ham <- subset(sms_raw, type == "ham")
wordcloud(spam$text, max.words = 50, scale = c(3, 0.5))
```

```{r}
wordcloud(ham$text, max.words = 50, scale = c(3, 0.5))
```

@@

```{r}
library(e1071)

sms_classifier <- naiveBayes(sms_train, sms_train_labels)
```

@@

```{r}
library(gmodels)
CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

As shown in the table only 29/1379 messages were classified incorrectly. This means that the algorithm classifed the testing set as spam or ham with approx. 98% accuracy. That’s impressive. To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.
I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.

@@

```{r}
wordcloud(spam$text, max.words = 5, scale = c(3, 0.5))

wordcloud(ham$text, max.words = 5, scale = c(3, 0.5))
```

@@

#My Test of Naive Bayes

##Loading Data Into R.
This block reads full CSV into a starting dataframe named "base_data". Next block isolates tables and rows into new dataframe "sms_raw2", and renames the columns.
```{r}
#CSV file named "coded_full_comments_dataset_Spring18anon.csv" already has correctly organized text and meta-data. First line reads it in.
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing <- filter(base_data, code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")
sms_raw2 <- frequency_writing %>% select(23,22)

#Rename columns
names(sms_raw2)[1] <- "type"
names(sms_raw2)[2] <- "text"
```

```{r}
str(sms_raw2)
```

The "type"" element is a character. This changes it to a factor for the analysis.

```{r}
sms_raw2$type <- factor(sms_raw2$type)
str(sms_raw2$type)
table(sms_raw2$type)
```

##Text Mining
For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using the "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.

```{r}
#install.packages("tm")
library(tm)
```

```{r}
sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
print(sms_corpus2)
```

@@

```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)
```

@@

@@

```{r}
#install.packages("SnowballC")
#library(SnowballC)

sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
```

@@

Because the dataset is random, the first 4180 messages can be used for the training set - there's no need to randomize the data first.

@@

```{r}
prop.table(table(sms_train_labels2))

prop.table(table(sms_test_labels2))
```

@@

```{}
install.packages("wordcloud")
install.packages("RColorBrewer")
```

@@

```{r}
TWO <- subset(sms_raw2, type == "2. Writing Quality")
THREE <- subset(sms_raw2, type == "3. Technical and Scientific")
FOUR <- subset(sms_raw2, type == "4. Logic and Thinking")

@@

The naive bayes classifier works with categorical reatures, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.

@@

```{r}
sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)

sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
```

@@

```{}
install.packages("e1071")
library(e1071)
```

```{r}
sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
```

```{r}
sms_test_pred2 <- predict(sms_classifier2, sms_test2)
```

```{r}
#library(gmodels)
CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

```{r}
1352/1379
```

As shown in the table only 27/1379 messages were classified incorrectly. This means that the algorithm classifed the testing set as spam or ham with approx. 98% accuracy. That's impressive. To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.

I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.

Show the 5 most frequent words in the sms data:

>>>>

#Demo of Naive Bayes (NB) Method
I copied the original demo web site [https://rpubs.com/Billyhansen6/318412](located here) to show how the process should work. I then duplicated the code and modified it to test with my TA comments.

The initial trial was able to identify sub-categories of comments ~80% of the time. This is better than the inter-rater reliability of two human raters, but still low. For subsequent iterations of the NB method I plan to evaluate:

*  Variations on data pre-processing: stemmed versus unstemmed, numbers removed vs. not, etc.
*  Comparing n-grams rather than single words
*  Pair-wise comparison versus comparing multiple groups simultaneously
*  Modifying the variables within the NB analysis itself.

####\

##Hansen's Original Version
The objective of this project is to classify SMS messages as spam or ham (not spam). A Naive Bayes classifier approach will be used. This example is taken from chapter 4 of Machine Learning with R, Second Edition"

@@

$$P(Spam|Hospital) = P(Hospital|Spam)P(Spam)/P(Hospital)$$

@@

####\

##Loading Libraries and Data

```{r}
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(tm)
library(SnowballC)
library(wordcloud)
library(e1071)
library(gmodels)
library(readr)
```

```{r}
#Demo dataset of SMS messages
sms_raw <- read.table("~/Dropbox/Coding_Tools/R_Environment/R_Projects/default_website/data/SMSSpamCollection.txt", header = FALSE, sep = "\t", quote = "", col.names = c("type","text"), stringsAsFactors = FALSE)
```

The data frame has 5574 observations marked as either spam or ham.

@@

Now the "type" variable is a factor with 2 levels. Of 5574 messages, 747 (about 13.4%) are spam.

####\

##Text Mining
For Naive Bayes to run effectively, the test data needs to be transformed. This begins with using "tm" to create a volitile coprus that contains the "text" vector from our data frame.

@@

```{r}
as.character(sms_corpus[[3]])
as.character((sms_corpus_clean[[3]]))
```

@@

```{r}
sms_corpus_clean <- tm_map(sms_corpus_clean, stemDocument)
```

@@

@@

```{r}
wordcloud(sms_corpus_clean, max.words = 50, random.order = FALSE)
```

Compare wordclouds between spam and ham.
```{r}
spam <- subset(sms_raw, type == "spam")
ham <- subset(sms_raw, type == "ham")
wordcloud(spam$text, max.words = 50, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 50, scale = c(3, 0.5))
```

####\

@@

```{r}
sms_classifier <- naiveBayes(sms_train, sms_train_labels)
```

####\

@@

```{r}
CrossTable(sms_test_pred, sms_test_labels, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

As shown in the table only 27/1379 messages were classified incorrectly. This means that the algorithm classifed the testing set as spam or ham with approx. 98% accuracy. That’s impressive. To improve the model, one might tamper with the Laplace value, colect more sms data, or try splitting the dataset randomly into training and testing.
I suspect the accuracy would increase as the dataset gets bigger. The more data there is to train the algorith, the more effective it would be in predicting Spam or Ham.

@@

```{r}
wordcloud(spam$text, max.words = 5, scale = c(3, 0.5))
wordcloud(ham$text, max.words = 5, scale = c(3, 0.5))
```

@@

#My Test of Naive Bayes
##Loading Data
This block reads full CSV into a starting dataframe named "base_data". Next block isolates tables and rows into new dataframe "sms_raw2", and renames the columns.
```{r}
#library(tm)
#library(SnowballC)
#library(wordcloud)
#library(e1071)
#library(gmodels)
#library(readr)
```

```{r}
#Read in CSV file named "coded_full_comments_dataset_Spring18anon.csv".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anon.csv')

#Isolate rows to compare. 
frequency_writing <- filter(base_data,code.subject=="2. Writing Quality"|code.subject=="3. Technical and Scientific"|code.subject=="4. Logic and Thinking")

#Reduce larger dataframe to 2 required columns of data, and put columns in order needed.
sms_raw2 <- frequency_writing %>% select(23,22)

#Rename the columns.
names(sms_raw2)[1] <- "type"
names(sms_raw2)[2] <- "text"

#Simplify the coding terms
sms_raw2[,1] <- ifelse(sms_raw2[,1] == "2. Writing Quality", "2_Writing", ifelse(sms_raw2[,1] == "3. Technical and Scientific", "3_Technical", ifelse(sms_raw2[,1] == "4. Logic and Thinking","4_Logic",99)))

#Change "type" element from character to a factor for analysis.
sms_raw2$type <- factor(sms_raw2$type)
str(sms_raw2$type)
table(sms_raw2$type)
```

####\

##Text Transformation Steps
This begins with using "tm" package in R to create a volitile coprus that contains the "text" vector from our data frame.

```{r}
#library(tm)
sms_corpus2 <- VCorpus(VectorSource(sms_raw2$text))
print(sms_corpus2)
```

@@

This removes numbers. **May not want to do.**
```{r}
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, removeNumbers)
```

Stopword removal takes out words that appear often but don't contribute to our objective. These words include "to", "and", "but" and "or". **One of variables to test out.**

@@

Remove punctuation as well using the "removePunctuation" function. **Try not doing this.**

@@

Perform "stemming" to the text data to strip the suffix from words. **Try NOT doing this.**

```{r}
#library(SnowballC)
sms_corpus_clean2 <- tm_map(sms_corpus_clean2, stemDocument)
```

@@

This code assumes comments are random. **Probably want to try randomizing them.**

@@

Make sure that the proportion of each sub-category is similar in the training and testing data set.

```{r}
prop.table(table(sms_train_labels2))
prop.table(table(sms_test_labels2))
```

####\

@@

@@

```{r}
TWO <- subset(sms_raw2, type == "2_Writing")
THREE <- subset(sms_raw2, type == "3_Technical")
FOUR <- subset(sms_raw2, type == "4_Logic")

@@

The naive bayes classifier works with categorical features, so we need to convert the matrix to "yes" and "no" categorical variables. To do this we'll build a convert_counts function and apply it to our data.

@@

```{r}
sms_train2 <- apply(sms_dtm_freq_train2, MARGIN = 2, convert_counts2)
sms_test2 <- apply(sms_dtm_freq_test2, MARGIN = 2, convert_counts2)
```

@@

####\

##Train Model, Predict, Evaluate
Use the e1071 package to implement Naive Bayes algorithm on the data, and predict whether a message is likely to be in group TWO, THREE, or FOUR. Evaluate the prediction with the actual data using a crosstable from the gmodels package.

```{r}
#library(e1071)
#library(gmodels)
sms_classifier2 <- naiveBayes(sms_train2, sms_train_labels2)
sms_test_pred2 <- predict(sms_classifier2, sms_test2)
CrossTable(sms_test_pred2, sms_test_labels2, prop.chisq = FALSE, prop.t = FALSE, dnn = c('predicted', 'actual'))
```

Show the 5 most frequent words in the data:
